\TheTitle

### 2020-01-04 Paper1/results.tex
last changes Martin, Ref NACA0012 missing

The aim is to quantify the effects arising from the one-dimensional uncertainty $\xi$ and to investigate its effects on the solution with different methods. To be able to measure the quality of the obtained solutions, we compute a reference solution using stochastic-Collocation with $100$ Gauss-Legendre quadrature points, which can be found in Figure~\ref{fig:referenceSolution}. In the following, we investigate the L$^2$-error of the variance and the expectation value. The L$^2$-error of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ is the cell average of the quantity $\bm e$ in spatial cell $j$, is denoted by
\begin{align*}
\Vert \bm e_{\Delta} \Vert_{\Delta} := \sqrt{\sum_{j=1}^{N_x} \Delta x_j \bm e_j^2}\;.
\end{align*}
Hence, when denoting the reference solution by $\bm u_{\Delta}$ and the moments obtained with the numerical method by $\bm{\hat u}_{\Delta}$, we investigate the relative error
\begin{align*}
\frac{\Vert \text{E}[\bm u_{\Delta}] - \text{E}[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert \text{E}[\bm u_{\Delta}] \Vert_{\Delta}} \qquad \text{ and }\qquad \frac{\Vert \text{Var}[\bm u_{\Delta}] - \text{Var}[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert \text{Var}[\bm u_{\Delta}] \Vert_{\Delta}}.
\end{align*}
The error is computed inside a box of one meter height and 1.1 meters length around the airfoil to prevent small fluctuations in the coarsely discretized far field from affecting the error.

@@

Let us now compare the results from intrusive methods with those of SC. For every quadrature point, SC iterates the solution until the density lies below a threshold of $\varepsilon = 1\cdot 10^{-7}$. Recording the error of intrusive methods during the iteration is straight forward. To record the error of SC, we couple all quadrature points after each iteration to evaluate the expectation value and variance, which destroys the black-box nature and results in additional costs. The collocation runtimes, that are depicted on the $x$-axis in Figures~\ref{fig:ErrorDifferentQuad}, \ref{fig:sub33}, \ref{fig:sub34} and \ref{fig:L2ErrorSolution2D}, are however rescaled to the runtimes that we achieve when running the collocation code in its original, black-box framework without recording the error. Our stochastic-Collocation computations use Clenshaw-Curtis quadrature levels $2,3$ and $4$, i.e. $5,9$ and $17$ quadrature points. The comparison of intrusive methods with non-intrusive methods shows the following:
\begin{itemize}
\item SC requires a smaller residual to converge to a steady state solution (we converge the results to $\varepsilon = 1\cdot 10^{-7}$ compared to $\varepsilon = 6\cdot 10^{-6}$ for intrusive methods).
\item Again, intrusive methods yield improved solutions compared to SC with the same number of unknowns. Actually, the error obtained with 17 unknowns when using SC is comparable with the error obtained with 10 unknowns when using intrusive methods.
\end{itemize}

@@

The error during the iteration has been recorded in Figure~\ref{fig:L2ErrorSolution2D}. To determine the error, we used stochastic-Collocation with a $50$ by $50$ Gauss-Legendre quadrature rule. As in the one-dimensional NACA testcase, the acceleration techniques lead to a heavily reduced runtime of the IPM method. Furthermore, the error obtained with the intrusive methods requires a total degree of $M = 9$, i.e. $N = 55$ moments to reach the error level of SC with 17 quadrature points per dimension i.e. $17^2 = 289$ collocation points. Note that this effect has also been observed in the one-dimensional case, however now for multi-dimensional problems, the reduced number of unknowns required for intrusive methods to obtain a certain error becomes more apparent. This shows one promising characteristic of intrusive techniques and points to their applicability for higher dimensional problems. In contrast to before, the SG error level is smaller than the IPM error for both, the expectation value and variance. Furthermore, when comparing IPM with and without One-Shot, one can observe that the effect of this acceleration strategy weighs in more heavily than it did for the one-dimensional case. This behavior is expected, since every iterate of the optimization problem becomes more expensive when the dimension is increased. This means, using a method with less iterations for every computation of the dual variables heavily reduces computational costs, which is why we consider the One-Shot strategy to be an effective approach, especially for problems with high uncertain dimension.

>>>>

The aim is to quantify the effects arising from the one-dimensional uncertainty $\xi$ and to investigate its effects on the solution with different methods. To be able to measure the quality of the obtained solutions, we compute a reference solution using Stochastic Collocation with $100$ Gauss-Legendre quadrature points, which can be found in Figure~\ref{fig:referenceSolution}. In the following, we investigate the L$^2$-error of the variance and the expectation value. The L$^2$-error of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ is the cell average of the quantity $\bm e$ in spatial cell $j$, is denoted by
\begin{align*}
\Vert \bm e_{\Delta} \Vert_{\Delta} := \sqrt{\sum_{j=1}^{N_x} \Delta x_j \bm e_j^2}\;.
\end{align*}
Hence, when denoting the reference solution by $\bm u_{\Delta}$ and the moments obtained with the numerical method by $\bm{\hat u}_{\Delta}$, we investigate the relative error
\begin{align*}
\frac{\Vert \text{E}[\bm u_{\Delta}] - \text{E}[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert \text{E}[\bm u_{\Delta}] \Vert_{\Delta}} \qquad \text{ and }\qquad \frac{\Vert \text{Var}[\bm u_{\Delta}] - \text{Var}[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert \text{Var}[\bm u_{\Delta}] \Vert_{\Delta}}.
\end{align*}
The error is computed inside a box of one meter height and 1.1 meters length around the airfoil to prevent small fluctuations in the coarsely discretized far field from affecting the error.

@@

Let us now compare the results from intrusive methods with those of SC. For every quadrature point, SC iterates the solution until the density lies below a threshold of $\varepsilon = 1\cdot 10^{-7}$. Recording the error of intrusive methods during the iteration is straight forward. To record the error of SC, we couple all quadrature points after each iteration to evaluate the expectation value and variance, which destroys the black-box nature and results in additional costs. The collocation runtimes, that are depicted on the $x$-axis in Figures~\ref{fig:ErrorDifferentQuad}, \ref{fig:sub33}, \ref{fig:sub34} and \ref{fig:L2ErrorSolution2D}, are however rescaled to the runtimes that we achieve when running the collocation code in its original, black-box framework without recording the error. Our Stochastic Collocation computations use Clenshaw-Curtis quadrature levels $2,3$ and $4$, i.e. $5,9$ and $17$ quadrature points. The comparison of intrusive methods with non-intrusive methods shows the following:
\begin{itemize}
\item SC requires a smaller residual to converge to a steady state solution (we converge the results to $\varepsilon = 1\cdot 10^{-7}$ compared to $\varepsilon = 6\cdot 10^{-6}$ for intrusive methods).
\item Again, intrusive methods yield improved solutions compared to SC with the same number of unknowns. Actually, the error obtained with 17 unknowns when using SC is comparable with the error obtained with 10 unknowns when using intrusive methods.
\end{itemize}

@@

The error during the iteration has been recorded in Figure~\ref{fig:L2ErrorSolution2D}. To determine the error, we used Stochastic Collocation with a $50$ by $50$ Gauss-Legendre quadrature rule. As in the one-dimensional NACA testcase, the acceleration techniques lead to a heavily reduced runtime of the IPM method. Furthermore, the error obtained with the intrusive methods requires a total degree of $M = 9$, i.e. $N = 55$ moments to reach the error level of SC with 17 quadrature points per dimension i.e. $17^2 = 289$ collocation points. Note that this effect has also been observed in the one-dimensional case, however now for multi-dimensional problems, the reduced number of unknowns required for intrusive methods to obtain a certain error becomes more apparent. This shows one promising characteristic of intrusive techniques and points to their applicability for higher dimensional problems. In contrast to before, the SG error level is smaller than the IPM error for both, the expectation value and variance. Furthermore, when comparing IPM with and without One-Shot, one can observe that the effect of this acceleration strategy weighs in more heavily than it did for the one-dimensional case. This behavior is expected, since every iterate of the optimization problem becomes more expensive when the dimension is increased. This means, using a method with less iterations for every computation of the dual variables heavily reduces computational costs, which is why we consider the One-Shot strategy to be an effective approach, especially for problems with high uncertain dimension.

### 2020-01-04 Paper1/summary_outlook.tex
last changes Martin, Ref NACA0012 missing

In future work, we aim at further accelerating the IPM method by using non-exact Hessian approximations. Similar to the One-Shot idea of not fully converging the dual problem, it seems to be plausible to not spend too much time on computing the Hessian when the moments are not close to a steady state. Hessian approximations that can be interesting are BFGS and sparse BFGS \cite[Chapter~6.1]{nocedal2006numerical}, which construct the Hessian from previously computed gradients. Note that this strategy will increase the used memory, since old Hessians or gradients from a certain number of old time steps need to be saved in every spacial cell.
Even though the non-intrusive nature of stochastic-Collocation or Monte Carlo methods allows an easy implementation, as shown in this work, it can be helpful to intrusively modify the code in order to fully exploit all acceleration potentials. Synchronizing the time updates of the solution at different quadrature points yields an increased control over the solution during the computation, which can for example be uses to employ adaptive methods. In this case one can switch to a fine quadrature level in a certain spatial cell by for example computing moments with the given coarse set of collocation points. From these moments one can compute an IPM reconstruction, which one can then evaluate at a finer quadrature set. Another example of breaking up the non-intrusive nature of Monte Carlo methods can be found in \cite{poette2019gpc}, where the generation of random samples is combined with the sampling after collisions to increase efficiency. Furthermore, we aim at applying the proposed acceleration techniques to SG methods with hyperbolicity limiters \cite{wu2017stochastic,schlachter2018hyperbolicity}.

>>>>

In future work, we aim at further accelerating the IPM method by using non-exact Hessian approximations. Similar to the One-Shot idea of not fully converging the dual problem, it seems to be plausible to not spend too much time on computing the Hessian when the moments are not close to a steady state. Hessian approximations that can be interesting are BFGS and sparse BFGS \cite[Chapter~6.1]{nocedal2006numerical}, which construct the Hessian from previously computed gradients. Note that this strategy will increase the used memory, since old Hessians or gradients from a certain number of old time steps need to be saved in every spacial cell.
Even though the non-intrusive nature of Stochastic Collocation or Monte Carlo methods allows an easy implementation, as shown in this work, it can be helpful to intrusively modify the code in order to fully exploit all acceleration potentials. Synchronizing the time updates of the solution at different quadrature points yields an increased control over the solution during the computation, which can for example be uses to employ adaptive methods. In this case one can switch to a fine quadrature level in a certain spatial cell by for example computing moments with the given coarse set of collocation points. From these moments one can compute an IPM reconstruction, which one can then evaluate at a finer quadrature set. Another example of breaking up the non-intrusive nature of Monte Carlo methods can be found in \cite{poette2019gpc}, where the generation of random samples is combined with the sampling after collisions to increase efficiency. Furthermore, we aim at applying the proposed acceleration techniques to SG methods with hyperbolicity limiters \cite{wu2017stochastic,schlachter2018hyperbolicity}.
